Accurate Binding of Sodium and Calcium to a POPC Bilayer by Effective Inclusion of Electronic Polarization

### 2018-03-15 Manuscript/SI.tex
Final changes Hector SI and Manuscript

\newpage
\section{Electronic continuum correction applied to CHARMM36 model}
In order to demonstrate that the ECC method can be applied also to other models than Lipid14,
we scaled the charges and Lennard-Jones radii of headgroup, glycerol backbone and
carbonyl atoms of the CHARMM36 model of POPC~\cite{klauda10}. We used the
scaling factors optimized for Lipid14 in the main text, i.e., 0.8 for charges and 0.89 for Lennard-Jones
radii. Changes of the headgroup order parameters as functions of added ions
and number density profiles along membrane normal from the ECC-CHARMM model are shown 
in Figs. \ref{scaledCHARMM} and \ref{scaledCHARMMdensity}, respectively,
together with the results from CHARMM36 simulations \cite{catte16,charmmPOPC450mMCaClfiles}
and experiments \cite{akutsu81,altenbach84}. As shown previously \cite{catte16},
the binding of Ca$^{2+}$ is significantly overestimated in CHARMM36 simulations.
In ECC-CHARMM simulations, the Ca$^{2+}$ binding affinity 
is significantly reduced (Fig. \ref{scaledCHARMMdensity}) and the headgroup order parameter response to the added CaCl$_2$
is in a better agreement with experiments than for the original CHARMM model (Fig. \ref{scaledCHARMM}).
In line with the results in main text,
Fig. \ref{scaledCHARMMcontacts} also shows faster equilibration times
for Ca$^{2+}$ binding in the ECC-CHARMM model than in the original CHARMM36.

In conclusion, applying ECC to CHARMM36 simulation
improved the ion binding affinity to a lipid bilayer, even thought the
scaling factors optimized for Lipid14 were used. Also the area per molecule in
the ECC-CHARMM simulation without ions (61 $\AA^2$) is only slightly
lower than the experimental value (64.3 $\AA^2$) \cite{kucerka11}. 
The optimal scaling factors for CHARMM36 are expected to be somewhat different
than for Lipid14 because the initial partial charges are not the same
in different models. Our results indicate that the performance of ion
binding in ECC-CHARMM36, and probably in other lipid models as well, could
be further improved by finding the optimal scaling factors for the partial
charges and Lennard-Jones radii.

It should be noted that the calcium binding to POPC bilayer within the ECC-CHARMM model
is compared here to the CHARMM36 simulation, which was originally done in a 
previous study \cite{catte16,charmmPOPC450mMCaClfiles}.
After the latter simulation has been performed, specific interaction parameters
between calcium and oxygen \cite{kim16} have been introduced in CHARMM-gui
to reduce the binding of calcium to lipid bilayers. These parameters
reduce the calcium binding to a POPC bilayer to the same level as that of sodium (data not shown).
Further analyses of ion binding with different parameters (including the newly modified CHARMM) are in progress
within the NMRlipids IV project (\url{nmrlipids.blogspot.fi}), being beyond the scope of this work.

>>>>

\newpage
\section{Electronic continuum correction applied to CHARMM36 model}
To demonstrate that the ECC method can also be applied to other models than Lipid14,
we scaled the charges and Lennard-Jones radii of headgroup, glycerol backbone and
carbonyl atoms of the CHARMM36 model of POPC~\cite{klauda10}. We used the
same scaling factors optimized for Lipid14 in the main text, i.e., 0.8 for charges and 0.89 for Lennard-Jones
radii. In Figs. \ref{scaledCHARMM} and \ref{scaledCHARMMdensity} we show
changes of the headgroup order parameters as functions of added ions
and number density profiles along membrane normal for our ECC-CHARMM model,
together with the CHARMM36 model \cite{catte16,charmmPOPC450mMCaClfiles}
and experiments \cite{akutsu81,altenbach84}. As shown previously \cite{catte16},
the binding of Ca$^{2+}$ is significantly overestimated in CHARMM36 simulations
where no explicit repulsive Ca$^{2+}$--oxygen Lennard-Jones parameters are included.
In ECC-CHARMM simulations, the Ca$^{2+}$ binding affinity 
is reduced substantially (Fig. \ref{scaledCHARMMdensity}), and the headgroup order parameter response to the added CaCl$_2$
is in a better agreement with experiments than for the original CHARMM36 model (Fig. \ref{scaledCHARMM}).
Fig. \ref{scaledCHARMMcontacts} also shows faster equilibration times
for Ca$^{2+}$ binding in the ECC-CHARMM model than in the original CHARMM36.
Both results are in line with the findings described in the main text.

In conclusion, applying ECC to CHARMM36 simulation
improved the ion binding affinity to a lipid bilayer, despite we
directly applied the scaling factors optimized for Lipid14 without further optimizations. Also, the area per molecule in
the ECC-CHARMM simulation without ions (61 $\AA^2$) is only slightly
lower than the experimental value (64.3 $\AA^2$) \cite{kucerka11}. 
The optimal scaling factors for CHARMM36 are expected to be somewhat different
than for Lipid14 because the initial partial charges are not the same
in different models. Our results indicate that the performance of ion
binding in ECC-CHARMM36, and probably in other lipid models as well, could
be further improved by finding the optimal values for the partial
charges and Lennard-Jones radii.

It should be noted that the CHARMM36 simulation used here uses the original paramter
as in our previous study \cite{catte16,charmmPOPC450mMCaClfiles}.
Currently (February 2018), specific interaction parameters
between calcium and oxygen \cite{kim16} have been introduced in CHARMM-GUI
to reduce the binding of calcium to lipid bilayers. The impact of this changes that
dramatically reduce the calcium binding to levels similar to sodium is beyond the
scope of this work.

### 2018-03-15 Manuscript/manuscript.tex
Final changes Hector SI and Manuscript

\subsubsection{Simulations of POPC bilayers with cationic surfactants}
The topology of the dihexadecyldimethylammonium cation was created with the automated topology builder \cite{malde11}.
The AmberTools program~\cite{amber} was used to generate the Amber-type force field parameters. These parameters were
then converted to the Gromacs format with the acpype tool \cite{acpype}. The partial charges were manually modified in
the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
For simulations of cations mixed with ECC-POPC, the ECC version of cation force field was made
by scaling the head group parameters with the same scaling factors that were
used for ECC-POPC, i.e., $f_q=0.8$ for charges and 0.89 for $\sigma$.

>>>>

\subsubsection{Simulations of POPC bilayers with cationic surfactants}
The topology of the dihexadecyldimethylammonium cation was created with the automated topology builder \cite{malde11}.
The AmberTools program~\cite{amber} was used to generate the Amber-type force field parameters. These parameters were
then converted to the Gromacs format with the acpype tool \cite{acpype}. The partial charges were manually modified in
the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
For simulations of cationic surfactant mixed with ECC-POPC, the ECC version of the cationic surgactnat force field was made
by scaling the partial charges and Lennard-Jones parameters of the head group parameters with the same scaling factors as in
ECC-POPC, i.e., $f_q=0.8$ for charges and 0.89 for $\sigma$.
